---
title: "MBAA 531 Test 4"
author: 
      - Brian McCormack, 568253728
      - Takashi Baba, 658190905
date: "12/10/2021"
output: html_document
---

### Setup
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
pokemonData <- read.csv(file = "data/pokemon.csv")
```

### There is substantial evidence that the pokemon HP data is normally distributed. The four main supporting arguments suggesting that this data follows a normal distribution are arguments 1.1, 1.2, 1.3, and 1.4
### 1.1 - Histogram of HP
#### The histogram below represents a nornal (bell shaped) distribution of data. Furthermore, the different colours in the graph refer to the different pokemon types. As shown the graph is symmetrical around a median HP of `r median(pokemonData$HP)`, represented by a vertical dashed line. Also the mean of `r mean(pokemonData$HP)`, represented by a solid vertical line is very close to the median
```{r}
hpHist <- ggplot(data = pokemonData, mapping = aes(x = HP, fill = Primary.Type)) + geom_histogram()
hpHist + geom_vline(mapping = aes(xintercept = mean(HP)), colour ='black') + geom_vline(mapping = aes(xintercept = median(HP)), linetype ="dashed", colour = 'black')
```

### 1.2 - Dot Plot of HP According to Pokemon Type
#### This dotplot shows that some pokemon types are more numerous than others. For instance Bug pokemon is numerous as the dots (cards) are distrubited in a normal fashion. In contrast, there are few flying pokemon. Therefore, there will be more variation in the flying pokemon due to a smaller dataset.
```{r}
ggplot(data = pokemonData, mapping = aes(x = HP, group = Primary.Type, fill = Primary.Type, )) + geom_dotplot(binwidth = NULL) + facet_wrap(~Primary.Type)
```

### 1.3 - Box Plot of HP
#### As shown below the 2nd and 3rd quartile are a similar distance from the median therefore the box in the box plot is mostly symmetrical.
```{r}
ggplot(data = pokemonData, mapping = aes(x = HP, y = 0)) + geom_jitter() + geom_boxplot(alpha  = .5)
```

### 1.4 - Normal Standard Deviation Rule
#### As shown below 73% of the points fall within one standard deviation, which is > 68% rule. In addtion 97% of the points fall within two standard deviations which is close to the 95% rule. This further suggests that the data is normally distributed. 
```{r}
#68 % rule
m = median(pokemonData$HP)
std = sd(pokemonData$HP)
index <- (pokemonData$HP >= m - std & pokemonData$HP <= m + std)
pointsInSd <- sum(index)/ length(index)
pointsInSd

#95% rule
index2 <- (pokemonData$HP >= m - 1.96*std & pokemonData$HP <= m + 1.96*std)
pointsInSd2 <- sum(index2) / length(index2)
pointsInSd2
```


