---
title: "mbaa531A4"
author: 
      - Brian McCormack
      - Takashi Baba
date: "12/10/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
pokemonData <- read.csv(file = "data/pokemon.csv")
```

## Pokemon HP Evaluation
### as shown below 

```{r}
#debugging
colnames(pokemonData)
head(pokemonData)
```
#### as shown above 

### histogram of hp
#### The histogram below represents a bell shaped distribution of data. Furthermore, the different colours in the graph refer to the different pokemon types. As shown the graph is symmetrical around a median HP of `r median(pokemonData$HP)`, represented by a vertical dashed line. Also the mean of `r mean(pokemonData$HP)`, represented by a solid vertical line is very close to the median
```{r}
hpHist <- ggplot(data = pokemonData, mapping = aes(x = HP, fill = Primary.Type)) + geom_histogram()
hpHist + geom_vline(mapping = aes(xintercept = mean(HP)), colour ='black') + geom_vline(mapping = aes(xintercept = median(HP)), linetype ="dashed", colour = 'black')
```

### dot plot of hp according to pokemon type
```{r}
ggplot(data = pokemonData, mapping = aes(x = HP, group = Primary.Type, fill = Primary.Type, )) + geom_dotplot(binwidth = NULL) + facet_wrap(~Primary.Type)
```

### box plot of hp
#### as shown below the 2nd and 3rd quartile are similar distance from the mean therefore the box in the box plot is mostly symmetrical
```{r}
ggplot(data = pokemonData, mapping = aes(x = HP, y = 0)) + geom_jitter() + geom_boxplot(alpha  = .5)
```

### averages
#### as shown below the 73% of the points fall within one standard deviation, which is > 68% rule suggesting that the data is normally distributed
```{r}
m = median(pokemonData$HP)
std = sd(pokemonData$HP)
index <- (pokemonData$HP >= m - std & pokemonData$HP <= m + std)
sum(index)/ length(index)
```


